---
title: "Prediction model using landmark analysis"
author: "Yibai Zhao, Chihyun Lee"
date: "2023-05-02"
output:
  rmdformats::robobook:
    thumbnails: no
    lightbox: true
    gallery: no
    highlight: tango
    use_bookdown: no
    toc_depth: 2
    fig_caption: yes
    code_folding: hide
    embed_fonts: no
    keep_md: false
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(ggplot2)
library(dplyr)
library(tidyr)
library(here)
library(survival)
library(survminer)
library(foreach)
library(doParallel)
library(gridExtra)
library(gtsummary)
registerDoParallel(8) 
```

# Load data

-   exclusion criteria:

    -   exclude subjects who didn't have biochemical recurrence (BCR), N=12495
    -   exclude subjects having BCR within 3 month after prostatectomy (RP), N=896
    -   exclude subjects having metastasis within 6 month of RP OR having metastasis before BCR, N=195
    -   exclude subjects having chemo therapy, N=10
    -   exclude subjects having adjuvant treatment (treatment within 6 month and before BCR), N=97

```{r}
data <- read.csv("/Volumes/home/MSKCC-biochemical-recurrence/data/data_2023-04-25.csv")
head(data)
```

# Data Dictionary

| Variable Name            | Definition                                                  | Notes                                       |
|-----------------|--------------------------------|-----------------------|
| `ptid`                   | Patient ID                                                  |                                             |
| `race`                   | Patient race                                                |                                             |
| `cci`                    | Charlson Comorbidity Index (CCI)                            |                                             |
| `sms`                    | Surgical margins                                            |                                             |
| `lni`                    | Lymph node involvement                                      |                                             |
| `dx_psa`                 | Last PSA before RP                                          |                                             |
| `tx_age`                 | Age at RP                                                   |                                             |
| `tx_year`                | Year of RP                                                  |                                             |
| `tx_type`                | RP type                                                     | RP / LP / RALP                              |
| `tx_grade`               | Pathological Gleason sum                                    |                                             |
| `tx_tstage`              | Pathological T stage                                        |                                             |
| `tt_bcr`                 | Time to BCR                                                 | 2 rises of \>=0.1 starting 42 days post RP  |
| `tt_calendar_bcr`        | Calendar year of time to BCR                                |                                             |
| `bcr_psa`                | PSA at BCR                                                  |                                             |
| `tt_salvage_tx`          | Time to salvage treatment                                   |                                             |
| `salvage_tx_type`        | Salvage treatment type                                      | Chemo / Hormones / RT / `NA` (No treatment) |
| `tt_calendar_salvage_tx` | Calendar year of time to salvage treatment                  |                                             |
| `tt_bcr_to_salvage_tx`   | Time (in month) from BCR to salvage treatment               |                                             |
| `outcome`                | Whether received salvage treatment within 6 month after BCR | 1-Received, 0-Not received                  |

: Data Dictionary

Note: Variable names have been renamed from original raw data set.

# Data preprocessing
```{r}
# 0. Filter out "Chemo" and patients having adjuvant treatment
data <- data %>% filter(salvage_tx_type != "Chemo" | is.na(salvage_tx_type))
data <- data %>% filter(is.na(tt_adjuvant_tx))
# 1. Categorization
data <- data %>%
  mutate(tx_grade = factor(tx_grade, levels = c("6", "3+4", "4+3", "8", "9", "10")))
data <- data %>% 
  mutate(tx_tstage_group = case_when(
    grepl("T2|T0", tx_tstage) ~ "<=T2",
    grepl("T3", tx_tstage) ~ "T3",
    grepl("T4", tx_tstage) ~ "T4"
  ))

data <- data %>%
  mutate(race_group = case_when(
    grepl("Asian|Other", race) ~ "Other",
    TRUE ~ race),
    # 2. Race="White" as baseline
    race_group = factor(race_group, levels = c("White", "Black", "Other")))

# 3. Reorder therapy factor
data <- data %>%
  mutate(across(salvage_tx_type, ~ifelse(is.na(tt_calendar_salvage_tx), "None", .)),
         salvage_tx_type = factor(salvage_tx_type, levels = c("None", "Hormones", "RT", "Hormones + RT")))
# age group
data <- data %>%
  mutate(tx_age_gp = cut(tx_age, c(0, 55, 70,100)))
# Categorize calendar year
data <- data %>%
  mutate(
    across(matches("calendar|year"), ~ifelse(.>= 2001 & .<=2003, "2001-2003",
                                        ifelse(.>= 2004 & .<=2006, "2004-2006",
                                               ifelse(.>= 2007 & .<=2009, "2007-2009",
                                                      ifelse(.>= 2010 & .<=2012, "2010-2012",
                                                             ifelse(.>= 2013 & .<=2015, "2013-2015",
                                                                    ifelse(.>= 2016 & .<=2018, "2016-2018",
                                                                           ifelse(.>= 2019 & .<=2021, "2019-2021", NA))))))),
           .names = "{col}_gp"),
    tt_calendar_bcr_gp2 = ifelse(tt_calendar_bcr >= 2001 & tt_calendar_bcr <= 2005, "2001-2005",
                                      ifelse(tt_calendar_bcr > 2005 & tt_calendar_bcr <= 2010, "2006-2010",
                                             ifelse(tt_calendar_bcr > 2010 & tt_calendar_bcr <= 2015, "2011-2015",
                                                    ifelse(tt_calendar_bcr > 2015 & tt_calendar_bcr <= 2021, "2016-2021", NA)))),
    across(matches("calendar|year"), ~factor(.)))
# if no salvage treatment, then tt_salvage_tx=tt_last_fu
data <- data %>% mutate(across(tt_salvage_tx, ~case_when(salvage_tx_type == "None" ~ tt_last_fu,
                                                         TRUE ~ .)))
# time(month) from bcr to salvage treatment
data <- data %>% mutate(tt_bcr_to_salvage_tx = tt_salvage_tx-tt_bcr)

# categorize comobility
data <- data %>%
  mutate(cci_gp = case_when(cci == 0 ~ "0",
                            cci <= 2 ~ "1-2",
                            TRUE ~ "3+")
         )

# indicator for immediate (trt within 6 mo) or late treatment (trt after 6 mo)
data <- data %>%
  mutate(salvage_tx_6mo = factor(case_when(tt_bcr_to_salvage_tx<=6 ~ "Immediate treatment",
                                    TRUE ~ "Late treatment")),
         salvage_tx_6mo = as.numeric(salvage_tx_6mo)-1) # 0-early trt, 1-late trt
# drop those died before receiving salvage treatment
# data <- data %>% filter(tt_death>tt_salvage_tx | is.na(tt_death))
#
# data %>% filter(salvage_tx_6mo==0) # 1009
# data %>% filter(salvage_tx_6mo==0 & tt_mets-tt_bcr<=6) # 51
# data %>% filter(salvage_tx_6mo==0 & tt_death-tt_bcr<=6) # 4 other death
# data %>% filter(salvage_tx_6mo==0 & tt_mets<tt_salvage_tx) # 37
# summary(data$tt_bcr_to_salvage_tx[data$salvage_tx_6mo==0])

# drop ptid=46427 due to 93-year of follow-up from RP
data <- data %>% filter(ptid!=46427)
```


# Landmark Analysis
Inclusion: both immediate and late treatment groups.

Event status: 0-censoring; 1-death
Event status2: 0-censoring; 2-metastasis
```{r}
# replace NA with Inf
data_landmk <- data %>%
  mutate(tt_bcr_yr = tt_bcr/12,
         tt_bcr_to_salvage_tx_yr = (tt_bcr_to_salvage_tx)/12,
         tt_bcr_mets_yr = (tt_mets-tt_bcr)/12,
         tt_bcr_death_yr = (tt_death-tt_bcr)/12,
         tt_bcr_last_fu_yr = (tt_last_fu-tt_bcr)/12) %>% 
  mutate_at(vars(tt_bcr_mets_yr, tt_bcr_death_yr, tt_bcr_last_fu_yr), ~replace_na(., Inf))
# event time and status
data_landmk <- data_landmk %>%
         mutate(tt_event_death = ifelse(is.infinite(tt_bcr_death_yr), tt_bcr_last_fu_yr, tt_bcr_death_yr),
                tt_event_mets = ifelse(is.infinite(tt_bcr_mets_yr), tt_bcr_last_fu_yr, tt_bcr_mets_yr),
                status_event_death = as.numeric(!is.infinite(tt_bcr_death_yr)),
                status_event_PCa = status_event_death==1 & death_type=="Prostate Cancer Death",
                status_event_other = status_event_death==1 & death_type=="Other Cause Death",
                status_event_mets = as.numeric(!is.infinite(tt_bcr_mets_yr))) # double check with Lukas

           # tt_event_death = ifelse(tt_bcr_last_fu_yr<tt_bcr_death_yr, tt_bcr_last_fu_yr, tt_bcr_death_yr),
           #      tt_event_mets = ifelse(tt_bcr_last_fu_yr<tt_bcr_mets_yr, tt_bcr_last_fu_yr, tt_bcr_mets_yr),
           #      status_event_death = as.numeric(!is.infinite(tt_bcr_death_yr)&(tt_bcr_last_fu_yr>=tt_bcr_death_yr)),
           #      status_event_mets = as.numeric(!is.infinite(tt_bcr_mets_yr)&(tt_bcr_last_fu_yr>=tt_bcr_mets_yr))) 
# data_landmk %>% select(ptid, tt_bcr_death_yr, tt_bcr_last_fu_yr, tt_event_death, status_event_death, tt_bcr_to_salvage_tx_yr, salvage_tx_type) %>% View
# indicator of salvage treatment
data_landmk <- data_landmk %>%
  mutate(ind_salvage_trt = ifelse(salvage_tx_type!="None"&!is.na(salvage_tx_type), 1, 0),
         ind_salvage_trt_6mo = ifelse((ind_salvage_trt==1)&(tt_bcr_to_salvage_tx_yr<0.5), 1, 0))
```

<!-- ## Censoring before death -->
<!-- ```{r} -->
<!-- data_landmk %>% filter(tt_bcr_last_fu_yr<tt_bcr_death_yr & !is.infinite(tt_bcr_death_yr)) %>% select(ptid, tt_bcr_to_salvage_tx_yr, salvage_tx_type, tt_bcr_last_fu_yr, tt_bcr_death_yr) -->
<!-- # censoring before having treatment and died after (ignoring) -->
<!-- data_landmk %>% filter(!is.infinite(tt_bcr_death_yr) & tt_bcr_last_fu_yr<tt_bcr_death_yr & salvage_tx_type=="None") %>% select(ptid, tt_bcr_to_salvage_tx_yr, salvage_tx_type, tt_bcr_last_fu_yr, tt_bcr_death_yr) -->
<!-- # having treatment before censoring and died after -->
<!-- data_landmk %>% filter(tt_bcr_last_fu_yr<tt_bcr_death_yr & salvage_tx_type!="None" & !is.infinite(tt_bcr_death_yr)) %>% select(ptid, tt_bcr_to_salvage_tx_yr, salvage_tx_type, tt_bcr_last_fu_yr, tt_bcr_death_yr) -->
<!-- ``` -->



## Propensity score and IPW
```{r}
data_landmk <- data_landmk %>%
  mutate(age_bcr=tx_age+tt_bcr/12)
data_landmk <- data_landmk %>% drop_na(age_bcr,bmi,cci,tx_grade,tx_tstage_group,sms,lni,bcr_psa,dx_psa,tt_bcr_yr)
# logistic regression to find significant covariates, covariates with NA
fit_logit <- glm(ind_salvage_trt_6mo~age_bcr+bmi+cci+tx_grade+tx_tstage_group+sms+lni+bcr_psa+dx_psa+tt_bcr_yr, family = binomial(link='logit'), data = data_landmk)
# fit_logit %>%
#    tbl_regression(exponentiate = TRUE)
# find propensity score
data_landmk <- data_landmk %>%
  mutate(glm_probs = predict(fit_logit, type="response"),
         weights = ifelse(ind_salvage_trt_6mo==1, 1/glm_probs, 1/(1-glm_probs)))
# Trimming approach to deal with extreme cases (PS close to 0 or 1)
glm_probs_quantile <- quantile(data_landmk$glm_probs, c(.025, .975))
data_landmk <- data_landmk %>% 
  mutate(glm_probs = ifelse(glm_probs<glm_probs_quantile[1], glm_probs_quantile[1], glm_probs),
         glm_probs = ifelse(glm_probs>glm_probs_quantile[2], glm_probs_quantile[2], glm_probs))
# data_landmk %>% select(ptid, glm_probs, weights) %>% View
```

https://www.emilyzabor.com/tutorials/survival_analysis_in_r_tutorial.html#Part_2:_Landmark_Analysis_and_Time_Dependent_Covariates
```{r}
# Subset population for those followed at least until landmark time
data_landmk_0.5 <- 
  data_landmk %>% 
  filter(tt_event_death > 0.5) 

# Calculate follow-up time from landmark and apply traditional methods
data_landmk_0.5 <- 
  data_landmk_0.5 %>% 
  mutate(
    tt_event_death = tt_event_death - 0.5,
    tt_bcr_to_salvage_tx_yr = tt_bcr_to_salvage_tx_yr - 0.5
    )

```
There are `r data_landmk %>% filter(tt_event_death <= 0.5) %>% count %>% as.numeric()` subjects who censored or died whithin 6-month after BCR.

## Time to all death
### KM plot
```{r}
library(ggsurvfit)
survfit(Surv(tt_event_death, status_event_death) ~ ind_salvage_trt_6mo, data = data_landmk_0.5) %>% 
  ggsurvfit() +
  labs(
    x = "Years from 6-month landmark",
    y = "Overall survival probability"
  ) +
  add_risktable()
survfit(Surv(tt_event_death, status_event_death) ~ ind_salvage_trt, data = data_landmk_0.5) %>% 
  ggsurvfit() +
  labs(
    x = "Years from 6-month landmark",
    y = "Overall survival probability"
  ) +
  add_risktable()
```

### Cox PH
```{r}
coxph(Surv(tt_event_death, status_event_death) ~ ind_salvage_trt_6mo, data = data_landmk_0.5) %>%
  tbl_regression(exponentiate = TRUE)
coxph(Surv(tt_event_death, status_event_death) ~ ind_salvage_trt, data = data_landmk_0.5) %>%
  tbl_regression(exponentiate = TRUE)
```

Adding IPW weights to Cox PH
```{r}
coxph(Surv(tt_event_death, status_event_death) ~ ind_salvage_trt_6mo, 
      weights = weights,
      data = data_landmk_0.5) %>%
  tbl_regression(exponentiate = TRUE)
coxph(Surv(tt_event_death, status_event_death) ~ ind_salvage_trt, 
      weights = weights,
      data = data_landmk_0.5) %>%
  tbl_regression(exponentiate = TRUE)
```

Incorperate time-dependent covariate

```{r}
# create dataset with time-dependent covariates
covs <- c("age_bcr", "cci", "tx_grade", "tx_tstage_group", "sms", "lni", "bcr_psa", "dx_psa", "tt_bcr_yr", "weights")
td_dat <- 
  tmerge(
    data1 = data_landmk_0.5 %>% select(ptid, tt_event_death, status_event_death, ind_salvage_trt_6mo, all_of(covs)), 
    data2 = data_landmk_0.5 %>% select(ptid, tt_event_death, status_event_death, tt_bcr_to_salvage_tx_yr, ind_salvage_trt, ind_salvage_trt_6mo, all_of(covs)), 
    id = ptid, 
    death = event(tt_event_death, status_event_death),
    salvage_tx = tdc(tt_bcr_to_salvage_tx_yr)
    )
coxph(
  Surv(time = tstart, time2 = tstop, event = death) ~ salvage_tx+ind_salvage_trt_6mo, #cluster = ptid,
  data = td_dat,
  ) %>% 
  tbl_regression(exp = TRUE)
coxph(
  Surv(time = tstart, time2 = tstop, event = death) ~ salvage_tx+ind_salvage_trt_6mo+
    age_bcr+cci+tx_grade+tx_tstage_group+sms+lni+bcr_psa+dx_psa+tt_bcr_yr, #cluster = ptid,
  data = td_dat,
  ) %>% 
  tbl_regression(exp = TRUE)
```

## Time to Prostate Cancer Death
### KM plot
```{r}
library(ggsurvfit)
survfit(Surv(tt_event_death, status_event_PCa) ~ ind_salvage_trt_6mo, data = data_landmk_0.5) %>% 
  ggsurvfit() +
  labs(
    x = "Years from 6-month landmark",
    y = "Overall survival probability"
  ) +
  add_risktable()
survfit(Surv(tt_event_death, status_event_PCa) ~ ind_salvage_trt, data = data_landmk_0.5) %>% 
  ggsurvfit() +
  labs(
    x = "Years from 6-month landmark",
    y = "Overall survival probability"
  ) +
  add_risktable()
```

### Cox PH
```{r}
coxph(Surv(tt_event_death, status_event_PCa) ~ ind_salvage_trt_6mo, data = data_landmk_0.5) %>%
  tbl_regression(exponentiate = TRUE)
coxph(Surv(tt_event_death, status_event_PCa) ~ ind_salvage_trt, data = data_landmk_0.5) %>%
  tbl_regression(exponentiate = TRUE)
```

Adding IPW weights to Cox PH
```{r}
coxph(Surv(tt_event_death, status_event_PCa) ~ ind_salvage_trt_6mo, 
      weights = weights,
      data = data_landmk_0.5) %>%
  tbl_regression(exponentiate = TRUE)
coxph(Surv(tt_event_death, status_event_PCa) ~ ind_salvage_trt, 
      weights = weights,
      data = data_landmk_0.5) %>%
  tbl_regression(exponentiate = TRUE)
```

Incorporate time-dependent covariates

```{r}
# create dataset with time-dependent covariates
covs <- c("age_bcr", "cci", "tx_grade", "tx_tstage_group", "sms", "lni", "bcr_psa", "dx_psa", "tt_bcr_yr", "weights")
td_dat <- 
  tmerge(
    data1 = data_landmk_0.5 %>% select(ptid, tt_event_death, status_event_PCa, ind_salvage_trt_6mo, all_of(covs)), 
    data2 = data_landmk_0.5 %>% select(ptid, tt_event_death, status_event_PCa, tt_bcr_to_salvage_tx_yr, ind_salvage_trt, ind_salvage_trt_6mo, all_of(covs)), 
    id = ptid, 
    death = event(tt_event_death, status_event_PCa),
    salvage_tx = tdc(tt_bcr_to_salvage_tx_yr)
    )
coxph(
  Surv(time = tstart, time2 = tstop, event = death) ~ salvage_tx+ind_salvage_trt_6mo, #cluster = ptid,
  data = td_dat,
  ) %>% 
  tbl_regression(exp = TRUE)
coxph(
  Surv(time = tstart, time2 = tstop, event = death) ~ salvage_tx+ind_salvage_trt_6mo+
    age_bcr+cci+tx_grade+tx_tstage_group+sms+lni+bcr_psa+dx_psa+tt_bcr_yr, #cluster = ptid,
  data = td_dat,
  ) %>% 
  tbl_regression(exp = TRUE)
```
```{r}
coxph(
  Surv(time = tt_event_death, event = status_event_PCa) ~ tt(tt_bcr_to_salvage_tx_yr)+ind_salvage_trt+ind_salvage_trt_6mo, 
  data = data_landmk_0.5,
  ) %>% 
  tbl_regression(exp = TRUE)
coxph(
  Surv(time = tt_event_death, event = status_event_PCa) ~ tt(tt_bcr_to_salvage_tx_yr)+ind_salvage_trt_6mo, 
  data = data_landmk_0.5,
  ) %>% 
  tbl_regression(exp = TRUE)

```


### Predict survival
```{r}
# time to treatment
fit_coxph <- coxph(
  Surv(time = tstart, time2 = tstop, event = death) ~ salvage_tx+ind_salvage_trt_6mo, 
  data = td_dat,
  ) 
ptid <- 10358
newdat <- td_dat %>% filter(ptid == !!ptid)
ggsurvplot(survfit(fit_coxph, newdata = newdata, id = ptid), data = newdata)

ptid <- 10358
newdata1 <- data_landmk_0.5 %>% filter(ptid == !!ptid)
newdata2 <- newdata1
newdata2$ind_salvage_trt <- 0
newdata2$ind_salvage_trt_6mo <- 0
newdata2$ptid <- newdata1$ptid+10
newdata <- rbind(newdata1, newdata2)
newdata_td <- 
  tmerge(
    data1 = newdata %>% select(ptid, tt_event_death, status_event_PCa, ind_salvage_trt_6mo, all_of(covs)), 
    data2 = newdata %>% select(ptid, tt_event_death, status_event_PCa, tt_bcr_to_salvage_tx_yr, ind_salvage_trt, ind_salvage_trt_6mo, all_of(covs)), 
    id = ptid, 
    death = event(tt_event_death, status_event_PCa),
    salvage_tx = tdc(tt_bcr_to_salvage_tx_yr)
    )
ggsurvplot(survfit(fit_coxph, newdata = newdata_td, id = ptid), data = newdata_td)

```

## Time to Other Cause Death
### KM plot
```{r}
library(ggsurvfit)
survfit(Surv(tt_event_death, status_event_other) ~ ind_salvage_trt_6mo, data = data_landmk_0.5) %>% 
  ggsurvfit() +
  labs(
    x = "Years from 6-month landmark",
    y = "Overall survival probability"
  ) +
  add_risktable()
survfit(Surv(tt_event_death, status_event_other) ~ ind_salvage_trt, data = data_landmk_0.5) %>% 
  ggsurvfit() +
  labs(
    x = "Years from 6-month landmark",
    y = "Overall survival probability"
  ) +
  add_risktable()
```

### Cox PH
```{r}
coxph(Surv(tt_event_death, status_event_other) ~ ind_salvage_trt_6mo, data = data_landmk_0.5) %>%
  tbl_regression(exponentiate = TRUE)
coxph(Surv(tt_event_death, status_event_other) ~ ind_salvage_trt, data = data_landmk_0.5) %>%
  tbl_regression(exponentiate = TRUE)
```

Adding IPW weights to Cox PH
```{r}
coxph(Surv(tt_event_death, status_event_other) ~ ind_salvage_trt_6mo, 
      weights = weights,
      data = data_landmk_0.5) %>%
  tbl_regression(exponentiate = TRUE)
coxph(Surv(tt_event_death, status_event_other) ~ ind_salvage_trt, 
      weights = weights,
      data = data_landmk_0.5) %>%
  tbl_regression(exponentiate = TRUE)
```


## Time to Metastasis
```{r}
# Subset population for those followed at least until landmark time
data_landmk_mets_0.5 <- 
  data_landmk %>% 
  filter(tt_event_mets > 0.5 & tt_event_death > 0.5) 

# Calculate follow-up time from landmark and apply traditional methods
data_landmk_mets_0.5 <- 
  data_landmk_mets_0.5 %>% 
  mutate(
    tt_event_mets = tt_event_mets - 0.5,
    tt_bcr_to_salvage_tx_yr = tt_bcr_to_salvage_tx_yr - 0.5
    )

```

There are `r data_landmk %>% filter(tt_event_mets <= 0.5 | tt_event_death <= 0.5) %>% count %>% as.numeric()` subjects who censored, died or had metastasis whithin 6-month after BCR.

### KM plot
```{r}
library(ggsurvfit)
survfit(Surv(tt_event_mets, status_event_mets) ~ ind_salvage_trt_6mo, data = data_landmk_mets_0.5) %>% 
  ggsurvfit() +
  labs(
    x = "Years from 6-month landmark",
    y = "Overall survival probability"
  ) +
  add_risktable()
survfit(Surv(tt_event_mets, status_event_mets) ~ ind_salvage_trt, data = data_landmk_mets_0.5) %>% 
  ggsurvfit() +
  labs(
    x = "Years from 6-month landmark",
    y = "Overall survival probability"
  ) +
  add_risktable()
```

### Cox PH
```{r}
coxph(Surv(tt_event_mets, status_event_mets) ~ ind_salvage_trt_6mo, data = data_landmk_mets_0.5) %>%
  tbl_regression(exponentiate = TRUE)
coxph(Surv(tt_event_mets, status_event_mets) ~ ind_salvage_trt, data = data_landmk_mets_0.5) %>%
  tbl_regression(exponentiate = TRUE)
```

Adding IPW weights to Cox PH
```{r}
coxph(Surv(tt_event_mets, status_event_mets) ~ ind_salvage_trt_6mo, 
      weights = weights,
      data = data_landmk_mets_0.5) %>%
  tbl_regression(exponentiate = TRUE)
coxph(Surv(tt_event_mets, status_event_mets) ~ ind_salvage_trt, 
      weights = weights,
      data = data_landmk_mets_0.5) %>%
  tbl_regression(exponentiate = TRUE)
```

Incorperate time-dependent covariates

```{r}
# create dataset with time-dependent covariates
covs <- c("age_bcr", "cci", "tx_grade", "tx_tstage_group", "sms", "lni", "bcr_psa", "dx_psa", "tt_bcr_yr", "weights")
td_dat <- 
  tmerge(
    data1 = data_landmk_mets_0.5 %>% select(ptid, tt_event_mets, status_event_mets, ind_salvage_trt_6mo, all_of(covs)), 
    data2 = data_landmk_mets_0.5 %>% select(ptid, tt_event_mets, status_event_mets, tt_bcr_to_salvage_tx_yr, ind_salvage_trt, ind_salvage_trt_6mo, all_of(covs)), 
    id = ptid, 
    mets = event(tt_event_mets, status_event_mets),
    salvage_tx = tdc(tt_bcr_to_salvage_tx_yr)
    )
coxph(
  Surv(time = tstart, time2 = tstop, event = mets) ~ salvage_tx+ind_salvage_trt_6mo, #cluster = ptid,
  data = td_dat,
  ) %>% 
  tbl_regression(exp = TRUE)
coxph(
  Surv(time = tstart, time2 = tstop, event = mets) ~ salvage_tx+ind_salvage_trt_6mo+
    age_bcr+cci+tx_grade+tx_tstage_group+sms+lni+bcr_psa+dx_psa+tt_bcr_yr, #cluster = ptid,
  data = td_dat,
  ) %>% 
  tbl_regression(exp = TRUE)
```

Patient I
```{r}
ptid <- 10358
```

