---
title: "Prediction model at baseline"
author: "Yibai Zhao"
date: "2023-03-07"
output:
  rmdformats::robobook:
    thumbnails: no
    lightbox: true
    gallery: no
    highlight: tango
    use_bookdown: no
    toc_depth: 2
    fig_caption: yes
    code_folding: hide
    embed_fonts: no
    keep_md: false
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(ggplot2)
library(dplyr)
library(tidyr)
library(here)
library(survival)
library(survminer)
library(foreach)
library(doParallel)
library(gridExtra)
library(gtsummary)
registerDoParallel(8) 
```

# Load data

-   Exclusion/inclusion criteria:

    -   include subjects who have biochemical recurrence (BCR) at least 3 month after prostatectomy (RP)
    -   exclude subjects who have metastasis before BCR
    -   exclude patients having chemo therapy, N=7
    -   exclude patients having adjuvant treatment (treatment within 6 month and before BCR), N=78
    -   exclude patients receiving salvage treatment earlier than BCR, N=~55

```{r}
data <- read.csv("/Volumes/home/MSKCC-biochemical-recurrence/data/data_2023-01-13.csv")
head(data)
```

# Data Dictionary

| Variable Name            | Definition                                                  | Notes                                       |
|-----------------|--------------------------------|-----------------------|
| `ptid`                   | Patient ID                                                  |                                             |
| `race`                   | Patient race                                                |                                             |
| `cci`                    | Charlson Comorbidity Index (CCI)                            |                                             |
| `sms`                    | Surgical margins                                            |                                             |
| `lni`                    | Lymph node involvement                                      |                                             |
| `dx_psa`                 | Last PSA before RP                                          |                                             |
| `tx_age`                 | Age at RP                                                   |                                             |
| `tx_year`                | Year of RP                                                  |                                             |
| `tx_type`                | RP type                                                     | RP / LP / RALP                              |
| `tx_grade`               | Pathological Gleason sum                                    |                                             |
| `tx_tstage`              | Pathological T stage                                        |                                             |
| `tt_bcr`                 | Time to BCR                                                 | 2 rises of \>=0.1 starting 42 days post RP  |
| `tt_calendar_bcr`        | Calendar year of time to BCR                                |                                             |
| `bcr_psa`                | PSA at BCR                                                  |                                             |
| `tt_salvage_tx`          | Time to salvage treatment                                   |                                             |
| `salvage_tx_type`        | Salvage treatment type                                      | Chemo / Hormones / RT / `NA` (No treatment) |
| `tt_calendar_salvage_tx` | Calendar year of time to salvage treatment                  |                                             |
| `tt_bcr_to_salvage_tx`   | Time (in month) from BCR to salvage treatment               |                                             |
| `outcome`                | Whether received salvage treatment within 6 month after BCR | 1-Received, 0-Not received                  |

: Data Dictionary

Note: Variable names have been renamed from original raw data set.

# Data preprocessing
```{r}
# 0. Filter out "Chemo" and patients having adjuvant treatment
data <- data %>% filter(salvage_tx_type != "Chemo" | is.na(salvage_tx_type))
data <- data %>% filter(is.na(tt_adjuvant_tx))
# 1. Categorization
data <- data %>%
  mutate(tx_grade_gp = case_when(
    tx_grade == "6" ~ "<=6",
    tx_grade %in% c("8", "9", "10") ~ ">7",
    TRUE ~ tx_grade
  ),
  tx_grade_gp = factor(tx_grade_gp, levels = c("<=6", "3+4", "4+3", ">7")))

data <- data %>% 
  mutate(tx_tstage_gp = case_when(
    grepl("T2|T0", tx_tstage) ~ "<=T2",
    grepl("T3", tx_tstage) ~ "T3",
    grepl("T4", tx_tstage) ~ "T4"
  ))

data <- data %>%
  mutate(race_gp = case_when(
    grepl("Asian|Other", race) ~ "Other",
    TRUE ~ race),
    # 2. Race="White" as baseline
    race_gp = factor(race_gp, levels = c("White", "Black", "Other")))

# 3. Reorer therapy factor
data <- data %>%
  mutate(across(salvage_tx_type, ~ifelse(is.na(tt_calendar_salvage_tx), "None", .)),
         salvage_tx_type = factor(salvage_tx_type, levels = c("None", "Hormones", "RT", "Hormones + RT")))
# age group
data <- data %>%
  mutate(tx_age_gp = cut(tx_age, c(0, 55, 70,100)))
# Categorize calendar year
data <- data %>%
  mutate(
    across(matches("calendar|year"), ~ifelse(.>= 2001 & .<=2003, "2001-2003",
                                        ifelse(.>= 2004 & .<=2006, "2004-2006",
                                               ifelse(.>= 2007 & .<=2009, "2007-2009",
                                                      ifelse(.>= 2010 & .<=2012, "2010-2012",
                                                             ifelse(.>= 2013 & .<=2015, "2013-2015",
                                                                    ifelse(.>= 2016 & .<=2018, "2016-2018",
                                                                           ifelse(.>= 2019 & .<=2021, "2019-2021", NA))))))),
           .names = "{col}_gp"),
    tt_calendar_bcr_gp2 = ifelse(tt_calendar_bcr >= 2001 & tt_calendar_bcr <= 2005, "2001-2005",
                                      ifelse(tt_calendar_bcr > 2005 & tt_calendar_bcr <= 2010, "2006-2010",
                                             ifelse(tt_calendar_bcr > 2010 & tt_calendar_bcr <= 2015, "2011-2015",
                                                    ifelse(tt_calendar_bcr > 2015 & tt_calendar_bcr <= 2021, "2016-2021", NA)))),
    across(matches("calendar|year"), ~factor(.)))
# if no salvage treatment, then tt_salvage_tx=tt_last_fu
data <- data %>% mutate(across(tt_salvage_tx, ~case_when(salvage_tx_type == "None" ~ tt_last_fu,
                                                         TRUE ~ .)))
# time(month) from bcr to salvage treatment
data <- data %>% mutate(tt_bcr_to_salvage_tx = tt_salvage_tx-tt_bcr)

# categorize comobility
data <- data %>%
  mutate(cci_gp = case_when(cci == 0 ~ "0",
                            cci <= 2 ~ "1-2",
                            TRUE ~ "3+")
         )

# indicator for immediate (trt within 6 mo) or late treatment (trt after 6 mo)
data <- data %>%
  mutate(salvage_tx_6mo = factor(case_when(tt_bcr_to_salvage_tx<=6 ~ "Immediate treatment",
                                    TRUE ~ "Late treatment")),
         salvage_tx_6mo = as.numeric(salvage_tx_6mo)-1) # 0-early trt, 1-late trt
```

# Data Exploration

## Baseline Characteristics

```{r}
# install.packages("gtsummary")
library(gtsummary)
data %>%
  dplyr::select(race, bmi, cci_gp, tx_age, tx_grade_gp, tx_tstage_gp, sms, lni, ln_pos_count, salvage_tx_type, tt_bcr, tt_salvage_tx, tt_last_fu, tt_mets, tt_death, death_type, salvage_tx_6mo) %>%
  tbl_summary(by = salvage_tx_6mo, missing = "no") %>%
  add_n() %>%
  add_p() %>% 
  modify_header(label = "**Variable**") #%>%
  # bold_labels() %>%
  # as_kable_extra(format = "latex")
```

## Histogram of time from BCR to salvage treatment for those received treatment after 6mo.
```{r}
data %>%
  filter(tt_bcr_to_salvage_tx>6 & tt_bcr_to_salvage_tx<500) %>% # Delete: one patient has salvage trt 1000 month after BCR
  ggplot(aes(x=tt_bcr_to_salvage_tx/12)) + 
  geom_histogram() +
  xlab("Time (in year) from BCR to Salvage Treatment")
```

# Regression


## Time to metastasis and death by immediate treatment VS late treatment
```{r}
# time to metastasis
KM_fit_mets_tx <- survfit(Surv((tt_mets-tt_bcr)/12, !is.na(tt_mets))~salvage_tx_6mo, data = data)
ggsurvplot(KM_fit_mets_tx, 
           data = data, 
           legend = "right", 
           legend.labs = levels(data$salvage_tx_6mo),
           legend.title = "",
           xlab = "Time (in years) from BCR to metastasis")
# time to PCa death
KM_fit_death_pca_tx <- survfit(Surv((tt_death-tt_bcr)/12, death_type=="Prostate Cancer Death")~salvage_tx_6mo, data = data)
ggsurvplot(KM_fit_death_pca_tx, 
           data = data, 
           legend = "right", 
           legend.labs = levels(data$salvage_tx_6mo),
           legend.title = "",
           xlab = "Time (in years) from BCR to PCa death")
# time to Other death
KM_fit_death_other_tx <- survfit(Surv((tt_death-tt_bcr)/12, death_type=="Other Cause Death")~salvage_tx_6mo, data = data)
ggsurvplot(KM_fit_death_other_tx, 
           data = data, 
           legend = "right", 
           legend.labs = levels(data$salvage_tx_6mo),
           legend.title = "",
           xlab = "Time (in years) from BCR to Other death")

```

## Time to PCa death stratify by Gleason grade and T stage by immediate treatment VS late treatment

```{r}
# grade
data_KM <- data %>%
  drop_na(tx_grade_gp, tx_tstage_gp) %>%
  select(tt_death, tt_bcr, death_type, salvage_tx_6mo, tx_grade_gp, tx_tstage_gp) 
data_KM_grade <- data_KM %>% 
  filter(tx_grade_gp!="<=6")
# data_KM <- data_KM %>% mutate(tx_grade_gp = factor(tx_grade_gp))
KM_fit_death_pca_tx_grade <- survfit(Surv((tt_death-tt_bcr)/12, death_type=="Prostate Cancer Death")~salvage_tx_6mo, data = data_KM_grade)
ggsurvplot(KM_fit_death_pca_tx_grade, data_KM_grade, 
                     facet.by = "tx_grade_gp",
                     xlab = "Time (in years) from BCR to PCa death")
data_KM_grade <- data_KM %>% 
  filter(tx_grade_gp=="<=6")
# data_KM <- data_KM %>% mutate(tx_grade_gp = factor(tx_grade_gp))
KM_fit_death_pca_tx_grade <- survfit(Surv((tt_death-tt_bcr)/12, death_type=="Prostate Cancer Death")~salvage_tx_6mo, data = data_KM_grade)
ggsurvplot(KM_fit_death_pca_tx_grade, data_KM_grade,
           facet.by = "tx_grade_gp",
           xlab = "Time (in years) from BCR to PCa death")

# T stage
data_KM_tstage <- data_KM %>% 
  filter(tx_tstage_gp!="<=T2")
KM_fit_death_pca_tx_tstage <- survfit(Surv((tt_death-tt_bcr)/12, death_type=="Prostate Cancer Death")~salvage_tx_6mo, data = data_KM_tstage)
ggsurvplot(KM_fit_death_pca_tx_tstage, data_KM_tstage, 
                     facet.by = "tx_tstage_gp",
                     xlab = "Time (in years) from BCR to PCa death")
data_KM_tstage <- data_KM %>% 
  filter(tx_tstage_gp=="<=T2")
KM_fit_death_pca_tx_tstage <- survfit(Surv((tt_death-tt_bcr)/12, death_type=="Prostate Cancer Death")~salvage_tx_6mo, data = data_KM_tstage)
ggsurvplot(KM_fit_death_pca_tx_tstage, data_KM_tstage, 
                     facet.by = "tx_tstage_gp",
                     xlab = "Time (in years) from BCR to PCa death")

```

# Landmark Analysis
Inclusion: Anyone who didn't receive immediate salvage treatment.

Event status: 0-censoring; 1-receive salvage treatment; 2-metastasis; 3-death
```{r}
# filter only non-immediate treatment
data_non6mo <- data %>% filter(tt_bcr_to_salvage_tx>6)
# replace NA with Inf
data_non6mo <- data_non6mo %>%
  mutate(tt_bcr_yr = tt_bcr/12,
         tt_bcr_to_salvage_tx_yr = tt_bcr_to_salvage_tx/12,
         tt_bcr_mets_yr = tt_mets-tt_bcr/12,
         tt_bcr_death_yr = tt_death-tt_bcr/12,
         tt_bcr_last_fu_yr = tt_last_fu-tt_bcr/12) %>% 
  mutate_at(vars(tt_bcr_mets_yr, tt_bcr_death_yr, tt_bcr_last_fu_yr), ~replace_na(., Inf))
# event status
data_non6mo <- data_non6mo %>%
         mutate(event_status=case_when(salvage_tx_type!="None" ~ 1,
                                       (salvage_tx_type=="None"&tt_bcr_mets_yr<=tt_bcr_last_fu_yr&tt_bcr_mets_yr<tt_bcr_death_yr) ~ 2,
                                  (salvage_tx_type=="None"&tt_bcr_death_yr<=tt_bcr_last_fu_yr&tt_bcr_death_yr<tt_bcr_mets_yr) ~ 3,
                              (salvage_tx_type=="None"&tt_bcr_last_fu_yr<tt_bcr_mets_yr&tt_bcr_last_fu_yr< tt_bcr_death_yr) ~ 0))
# event time
data_non6mo <- data_non6mo %>% rowwise() %>% mutate(event_time=min(tt_bcr_to_salvage_tx_yr, tt_bcr_mets_yr, tt_bcr_death_yr, tt_bcr_last_fu_yr)) %>% unnest(event_time)
table(data_non6mo$event_status)
```

```{r}
library(Landmarking)
## load PSA data
data_psa <- readRDS("/Volumes/home/MSKCC-biochemical-recurrence/data/psa_test_data.rds")
data_psa$ptid <- as.numeric(data_psa$ptid)
attr(data_psa$psa, 'problems') <- NULL
data_psa$time <- data_psa$time/12
data_all <- data_non6mo %>% left_join(data_psa, by = 'ptid')
```


```{r}
## fit last observation carried forward (LOCF) landmark model
data_model_landmark_LOCF <-
  fit_LOCF_landmark( # last observational value at landmark time
    data_long = data_all,
    x_L = c(2), # landmark time
    x_hor = c(10), # survival time
    covariates = c("psa","tx_grade_gp"),
    covariates_time ="time",
    individual_id = "ptid",
    event_time = "event_time",
    event_status = "event_status",
    survival_submodel = "fine_gray")
data_model_landmark_LOCF
```
## Patient I
```{r}
# time to treatment
ptid <- 10358
landmark_time <- 2
survival_time <- 20
covariates = c("psa", "tx_age", "cci", "tx_grade_gp", "tx_tstage_gp", "sms", "lni", "tt_bcr_yr")
covariates_time = "time"
data_model_landmark_LOCF <- 
      fit_LOCF_landmark( # last observational value at landmark time
        data_long = data_all,
        x_L = landmark_time, # landmark time
        x_hor = survival_time, # survival time
        covariates = covariates,
        covariates_time = covariates_time,
        individual_id = "ptid",
        event_time = "event_time",
        event_status = "event_status",
        survival_submodel = "cause_specific")
predict_landmark_LOCF <- function(dset, fit, landmark_time=2, survival_time, covariates, covariates_time="time", ptid){
  newdata <- dset %>% filter(ptid == !!ptid) %>% select(ptid, all_of(c(covariates, covariates_time)))
  pred <-
    predict(
      object = fit,
      x_L = landmark_time,
      x_hor = survival_time,
      newdata = newdata,
    )
  pred
}

oset_all <- foreach (tt_surv = seq(2, 20, 0.5), .combine = rbind) %dopar% {
  oset <- predict_landmark_LOCF(dset = data_all, 
                        fit = data_model_landmark_LOCF,
                       survival_time = tt_surv, 
                       covariates = covariates, 
                       ptid = ptid)
  oset$surv_t <- tt_surv
  oset
}
tb_obs <- data_all %>% filter(ptid == !!ptid) %>% select(ptid, all_of(c(covariates_time, covariates))) %>% filter(time<=landmark_time)
tb_obs
# tt <- ttheme_default(colhead=list(fg_params = list(parse=TRUE)),
#                      base_size = 8,
#                      padding = unit(c(1, 10), "mm"))
# tb_obs <- tableGrob(tb_obs, rows=NULL, theme=tt)
ggplot(data = oset_all, aes(x = surv_t, y = event_prediction)) +
  geom_line() +
  ylim(0, 1) +
  xlab("Time (in Year) from BCR") +
  ylab("CIF of Salvage Treatment") +
  theme_bw()
# grid.arrange(gg1, tb_obs, 
#              nrow = 2, heights = c(2, 0.5))


# time to metastasis
data_all2 <- data_all
data_all2$event_status[data_all$event_status==2] <- 1
data_all2$event_status[data_all$event_status==1] <- 2
data_model_landmark_LOCF2 <- 
      fit_LOCF_landmark( # last observational value at landmark time
        data_long = data_all2,
        x_L = landmark_time, # landmark time
        x_hor = survival_time, # survival time
        covariates = covariates,
        covariates_time = covariates_time,
        individual_id = "ptid",
        event_time = "event_time",
        event_status = "event_status",
        survival_submodel = "cause_specific")
oset_all2 <- foreach (tt_surv = seq(2, 20, 0.5), .combine = rbind) %dopar% {
  oset2 <- predict_landmark_LOCF(dset = data_all2, 
                        fit = data_model_landmark_LOCF2,
                       survival_time = tt_surv, 
                       covariates = covariates, 
                       ptid = ptid)
  oset2$surv_t <- tt_surv
  oset2
}
# data_all2 %>% filter(ptid == !!ptid) %>% select(ptid, all_of(c(covariates_time, covariates))) %>% filter(time<=landmark_time)
# tb_obs2 <- tableGrob(tb_obs2, rows=NULL, theme=tt)
ggplot(data = oset_all2, aes(x = surv_t, y = event_prediction)) +
  geom_line() +
  ylim(0, 1) +
  xlab("Time (in Year) from BCR") +
  ylab("CIF of Metastasis")+
  theme_bw()

```

## Patient II
```{r}
# time to treatment
ptid <- 6256
landmark_time <- 2
survival_time <- 20
covariates = c("psa", "tx_age", "cci", "tx_grade_gp", "tx_tstage_gp", "sms", "lni", "tt_bcr_yr")
covariates_time = "time"
data_model_landmark_LOCF <- 
      fit_LOCF_landmark( # last observational value at landmark time
        data_long = data_all,
        x_L = landmark_time, # landmark time
        x_hor = survival_time, # survival time
        covariates = covariates,
        covariates_time = covariates_time,
        individual_id = "ptid",
        event_time = "event_time",
        event_status = "event_status",
        survival_submodel = "cause_specific")
predict_landmark_LOCF <- function(dset, fit, landmark_time=2, survival_time, covariates, covariates_time="time", ptid){
  newdata <- dset %>% filter(ptid == !!ptid) %>% select(ptid, all_of(c(covariates, covariates_time)))
  pred <-
    predict(
      object = fit,
      x_L = landmark_time,
      x_hor = survival_time,
      newdata = newdata,
    )
  pred
}

oset_all <- foreach (tt_surv = seq(2, 20, 0.5), .combine = rbind) %dopar% {
  oset <- predict_landmark_LOCF(dset = data_all, 
                        fit = data_model_landmark_LOCF,
                       survival_time = tt_surv, 
                       covariates = covariates, 
                       ptid = ptid)
  oset$surv_t <- tt_surv
  oset
}
data_all %>% filter(ptid == !!ptid) %>% select(ptid, all_of(c(covariates_time, covariates))) %>% filter(time<=landmark_time)
ggplot(data = oset_all, aes(x = surv_t, y = event_prediction)) +
  geom_line() +
  ylim(0, 1) +
  xlab("Time (in Year) from BCR") +
  ylab("CIF of Salvage Treatment") +
  theme_bw()


# time to metastasis
data_all2 <- data_all
data_all2$event_status[data_all$event_status==2] <- 1
data_all2$event_status[data_all$event_status==1] <- 2
data_model_landmark_LOCF2 <- 
      fit_LOCF_landmark( # last observational value at landmark time
        data_long = data_all2,
        x_L = landmark_time, # landmark time
        x_hor = survival_time, # survival time
        covariates = covariates,
        covariates_time = covariates_time,
        individual_id = "ptid",
        event_time = "event_time",
        event_status = "event_status",
        survival_submodel = "cause_specific")
oset_all2 <- foreach (tt_surv = seq(2, 20, 0.5), .combine = rbind) %dopar% {
  oset2 <- predict_landmark_LOCF(dset = data_all2, 
                        fit = data_model_landmark_LOCF2,
                       survival_time = tt_surv, 
                       covariates = covariates, 
                       ptid = ptid)
  oset2$surv_t <- tt_surv
  oset2
}
ggplot(data = oset_all2, aes(x = surv_t, y = event_prediction)) +
  geom_line() +
  ylim(0, 1) +
  xlab("Time (in Year) from BCR") +
  ylab("CIF of Metastasis")+
  theme_bw()


```

# Cox PH analysis with propensity score weighting
```{r}
# age at BCR
data <- data %>%
  mutate(age_bcr=tx_age+tt_bcr/12,
         tt_bcr_year=tt_bcr/12)
data_coxph <- data %>% drop_na(age_bcr,bmi,cci,tx_grade,tx_tstage_gp,sms,lni,bcr_psa,dx_psa,tt_bcr_year)
```

```{r}
# logistic regression to find significant covariates, covariates with NA
fit_logit <- glm(salvage_tx_6mo~age_bcr+bmi+cci+tx_grade+tx_tstage_gp+sms+lni+bcr_psa+dx_psa+tt_bcr_year, family = binomial(link='logit'), data = data_coxph)
fit_logit %>%
   tbl_regression(exponentiate = TRUE)
# find propensity score
data_coxph <- data_coxph %>%
  mutate(glm_probs = predict(fit_logit, type="response"),
         weights = ifelse(salvage_tx_6mo==0, 1/(1-glm_probs), 1/glm_probs))
data_coxph %>% select(ptid, glm_probs, weights) %>% View
# Trimming approach to deal with extreme cases (PS close to 0 or 1)
glm_probs_quantile <- quantile(data_coxph$glm_probs, c(.025, .975))
data_coxph <- data_coxph %>% 
  mutate(glm_probs = ifelse(glm_probs<glm_probs_quantile[1], glm_probs_quantile[1], glm_probs),
         glm_probs = ifelse(glm_probs>glm_probs_quantile[2], glm_probs_quantile[2], glm_probs))
```

```{r}
new_df <- with(data_coxph,
               data.frame(#salvage_tx_6mo = c("Immediate treatment", "Late treatment"), 
                 salvage_tx_6mo = c(0, 1), 
                          age_bcr = 65,
                          cci = 1,
                          tx_grade = "8",
                          tx_tstage_gp = "<=T2",
                          sms = "Negative",
                          lni = "Negative",
                          bcr_psa = 2,
                          dx_psa = 2,
                          tt_bcr_year = 1 # year
               )
)
```

## Prostate cancer death
```{r}
# coxph
fit_pca <- coxph(Surv((tt_death-tt_bcr)/12, death_type=='Prostate Cancer Death') ~ salvage_tx_6mo+age_bcr+cci+tx_grade+tx_tstage_gp+sms+lni+bcr_psa+dx_psa+tt_bcr_year, 
                   weights = weights, data = data_coxph)
fit_pca %>%
  gtsummary::tbl_regression(exponentiate = TRUE) 
library(survminer)
pred_pca <- survfit(fit_pca, newdata = new_df)
ggsurvplot(pred_pca, data = new_df, 
           legend.labs = c("Immediate treatment", "Late treatment"),
           xlab = "Time (in years) from BCR",
           title = "Prostate cancer death",
           ggtheme = theme_minimal())

fit_pca2 <- coxph(Surv((tt_death-tt_bcr)/12, death_type=='Prostate Cancer Death') ~ salvage_tx_6mo+age_bcr+cci+tx_grade+tx_tstage_gp+sms+lni+bcr_psa+dx_psa+tt_bcr_year,
                 data = data_coxph)
fit_pca2 %>%
  gtsummary::tbl_regression(exponentiate = TRUE) 
pred_pca2 <- survfit(fit_pca2, newdata = new_df)
ggsurvplot(pred_pca2, data = new_df, 
           legend.labs = c("Immediate treatment", "Late treatment"),
           xlab = "Time (in years) from BCR",
           title = "Prostate cancer death",
           ggtheme = theme_minimal())

```

## Metastasis
```{r}
# coxph
fit_mets <- coxph(Surv((tt_mets-tt_bcr)/12, !is.na(tt_mets)) ~ salvage_tx_6mo+age_bcr+cci+tx_grade+tx_tstage_gp+sms+lni+bcr_psa+dx_psa+tt_bcr_year, 
                   weights = weights, data = data_coxph)
fit_mets %>%
  gtsummary::tbl_regression(exponentiate = TRUE) 

library(survminer)
pred_mets <- survfit(fit_mets, newdata = new_df)
ggsurvplot(pred_mets, data = new_df, 
           legend.labs = c("Immediate treatment", "Late treatment"),
           xlab = "Time (in years) from BCR",
           title = "Metastasis",
           ggtheme = theme_minimal())

fit_mets2 <- coxph(Surv((tt_mets-tt_bcr)/12, !is.na(tt_mets)) ~ salvage_tx_6mo+age_bcr+cci+tx_grade+tx_tstage_gp+sms+lni+bcr_psa+dx_psa+tt_bcr_year, 
                   data = data_coxph)
fit_mets2 %>%
  gtsummary::tbl_regression(exponentiate = TRUE) 

library(survminer)
pred_mets2 <- survfit(fit_mets2, newdata = new_df)
ggsurvplot(pred_mets2, data = new_df, 
           legend.labs = c("Immediate treatment", "Late treatment"),
           xlab = "Time (in years) from BCR",
           title = "Metastasis",
           ggtheme = theme_minimal())

```
## Check balance of covariates after propensity score
```{r}
library(cobalt)
####### Results explanation
# By setting stats = c("m", "v", "ks") we requested mean differences, variance ratios, and Kolmogorov-Smirnov statistics. 
####### Results explanation
bal.tab(salvage_tx_6mo ~ age_bcr + bmi + cci + tx_grade + 
    tx_tstage_gp + sms + lni + bcr_psa + dx_psa + tt_bcr_year, data = data_coxph, 
    un = TRUE, stats = c("m", "v", "ks"),
    weights = data_coxph$weights)
# Standardize Mean Difference (SMD)
love.plot(salvage_tx_6mo ~ age_bcr + bmi + cci + tx_grade + 
    tx_tstage_gp + sms + lni + bcr_psa + dx_psa + tt_bcr_year, data = data_coxph, 
    weights = data_coxph$weights, binary = "std", un = TRUE)
```


# Questions:
## How to define last followup? Patients died before lost to followup.
```{r}
data %>% filter(tt_death<tt_last_fu) %>% select(ptid, tt_death, salvage_tx_type, tt_salvage_tx, tt_last_fu)
```


## How many patients died before receiving immediate treatment (6 month)
```{r}
# death, how to treat them? should they belong to immediate or late treatment group?
data %>% filter(tt_death-tt_bcr<6 & salvage_tx_type=="None") %>% select(ptid, tt_death, tt_bcr, salvage_tx_type, tt_salvage_tx, tt_last_fu)
# metastasis
data %>% filter(tt_mets-tt_bcr<6 & tt_mets<tt_salvage_tx) %>% select(ptid, tt_death, tt_mets, tt_bcr, salvage_tx_type, tt_salvage_tx, tt_last_fu)
```

## How many paitents died/metas before receiving late salvage treatment 
```{r}
# death
data %>% filter(tt_death-tt_bcr>=6 & salvage_tx_type=="None" & tt_death<=tt_last_fu) %>% select(ptid, tt_death, tt_bcr, salvage_tx_type, tt_salvage_tx, tt_last_fu)
# metastasis
data %>% filter(tt_mets-tt_bcr>=6 & tt_mets<tt_salvage_tx) %>% select(ptid, tt_death, tt_mets, tt_bcr, salvage_tx_type, tt_salvage_tx, tt_last_fu)

```

